---
title: "Sok-1005_Assignment_1"
author: Daniel Johannessen. kandidat nummer 3
format: html
editor: visual
---

## Assignment 1 Reading static webpages. 

We start by gathering the data we need and getting all our packages

```{r, warning=False}
rm(list=ls()) #emptying enviroment
library(tidyverse)
library(lubridate)


lower <- read_table("http://vortex.nsstc.uah.edu/data/msu/v6.0/tlt/uahncdc_lt_6.0.txt")
mid <- read_table("http://vortex.nsstc.uah.edu/data/msu/v6.0/tmt/uahncdc_mt_6.0.txt")
tropopause <- read_table("http://vortex.nsstc.uah.edu/data/msu/v6.0/ttp/uahncdc_tp_6.0.txt")
stratosphere <- read_table("http://vortex.nsstc.uah.edu/data/msu/v6.0/tls/uahncdc_ls_6.0.txt")
```

Now that we have gotten our data and our packages, i start with removing the last rows in the dataset that contains comments that we don't need.

Here we use code to dynamically remove the comments even if new data is added. The code is gathered from the link provided below.

[Source of the code used for dynamic code](https://github.com/uit-sok-1005-v23/uit-sok-1005-v23.github.io/blob/main/Global_Temperature.R "Line 43")

```{r}

lower <- lower[1:which(lower$Year %in% "Year")-1, ]
mid <- mid[1:which(mid$Year %in% "Year")-1, ]
tropopause <- tropopause[1:which(tropopause$Year %in% "Year")-1, ]
stratosphere <- stratosphere[1:which(stratosphere$Year %in% "Year")-1, ]


```

We are told to also only use data in the Globe column so I will now select it out of each dataset then combine the datasets into one.

```{r}
lower <- lower %>% 
  mutate(Date = ym(paste(Year, Mo))) %>% #Creating a Date variable
  select(Date , Year, Mo, Globe) #selecting and arranging the data we want

mid <- mid %>% 
  mutate(Date = ym(paste(Year, Mo))) %>% 
  select(Date ,Year, Mo, Globe) 

stratosphere <- stratosphere %>% 
  mutate(Date = ym(paste(Year, Mo))) %>% 
  select(Date, Year, Mo, Globe) 

tropopause <- tropopause %>% 
  mutate(Date = ym(paste(Year, Mo))) %>% 
  select(Date, Year, Mo, Globe) 

```
